---
title: "Economic growth and productivity"
execute:
  warning: false
  message: false
execute-dir: project
format: 
  html:
    embed-resources: true
    code-fold: true
    code-summary: "Show code"
---

DRAFT

Updated: `r Sys.Date()`

```{r}
#| label: setup
# library(fiprod)

if (interactive()) devtools::load_all(".") else library(fiprod) 

library(tidyverse)
library(ggcustom)
library(pttdatahaku)


set_gg(theme_fpb())


dat_oecd_pdb_main <- load_dat("dat_oecd_pdb_main") 

dat_oecd_pdb_ind <- load_dat("dat_oecd_pdb_ind") 

dat_gva_ind <- load_dat("dat_gva_ind") 

geos <- rev(c(
  "Finland"         = "FI",
  "Euro area"      = "EA20",
  "Sweden"        = "SE",
  "Denmark"        = "DK",
  "Germany"         = "DE",
  "USA"           = "US",
  "others"          = "Other"
))

geos_width <- set_names(if_else(geos %in% c("FI"), 2, 1.3), names(geos))
geos_colour <- set_names(c("grey80", rev(ggcustom_pal(length(geos) -1, "fpb"))), names(geos)) # grey for the others category


# geos_name <- set_names(names(geos), geos)

```

## Data on economic growth and productivity development

The main source used is the [OECD productivity database](https://data-explorer.oecd.org/vis?fs%5B0%5D=Topic%2C1%7CEconomy%23ECO%23%7CProductivity%23ECO_PRO%23&pg=0&fc=Topic&bp=true&snb=7&df%5Bds%5D=dsDisseminateFinalDMZ&df%5Bid%5D=DSD_PDB%40DF_PDB&df%5Bag%5D=OECD.SDD.TPS&df%5Bvs%5D=2.0&dq=.A.GVAHRS._T.XDC_H..N..&lom=LASTNPERIODS&lo=5&to%5BTIME_PERIOD%5D=false&isAvailabilityDisabled=false) data ([database revamp](https://www.oecd.org/en/publications/the-revamp-of-the-oecd-productivity-database_f07c55d4-en.html) and [metadata](https://www.oecd.org/content/dam/oecd/en/data/methods/OECD-Productivity-Statistics-Database-metadata.pdf)).

For industries, price level data come from the [GGDC Productivity level](https://www.rug.nl/ggdc/productivity/pld/) database.

## GDP growth

```{r}

# OECD codes:
# 
# V Current prices
# 
# LR base year 2020 (also L, where the base year varies).

dat_oecd_pdb_main |> 
  filter(time >= "1995-01-01") |> 
  filter_recode(
    geo = geos,
    measure = c("GDPPOP"),
    activity = c("Total economy" = "_T"),
    price_base = c("LR"),        # 
    conversion_type = c("PPP")
  ) |> 
  select(-unit_measure) |> 
  mutate(values = rebase(values, time, baseyear = 2007), .by = where(is.factor)) |>  
  ggplot(aes(time, values, colour = geo, linewidth = geo)) +
  geom_line() +
  scale_linewidth_manual(values = geos_width) +
  scale_colour_manual(values = geos_colour) +
  guides(colour = guide_legend(reverse = TRUE), 
         linewidth = guide_legend(reverse = TRUE)) +
  the_title_blank("xyl") +
  labs(
    title = "GDP per capita",
    subtitle = "Index, 2007 = 100",
    caption = "Source: OECD, Finnish Productivity Board"
  )
  
```

## Growth split into productivity and hours worked

```{r}

dat_oecd_pdb_main |> 
  filter(time >= "1995-01-01") |> 
  filter_recode(
    geo = geos,
    measure = c("Labour productivity" = "GVAHRS", "Hours per capita" = "HRSPOP"),
    activity = c("Total economy" = "_T"),
    unit_measure = c("USD_PPP_H", "H_PS"),
    price_base = c("LR", "_Z"),        
    conversion_type = c("PPP", "_Z")
  ) |> 
  select(-unit_measure) |> 
  mutate(values = rebase(values, time, baseyear = 2007), .by = where(is.factor)) |> 
  ggplot(aes(time, values, colour = geo, linewidth = geo)) +
  facet_wrap(~measure) +
  geom_line() +
  scale_colour_manual(values = geos_colour) +
  scale_linewidth_manual(values = geos_width) +
  guides(colour = guide_legend(reverse = TRUE), linewidth = guide_legend(reverse = TRUE)) +
  the_title_blank("xyl") +
  labs(
    title = "GDP per capita",
    subtitle = "Index, 2007 = 100",
    caption = "Source: OECD, Finnish Productivity Board"
  )
```

## Labour productivity growth in the private sector and elsewhere

```{r}

dat_gva_ind |> 
  filter(time >= "1995-01-01") |> 
  filter_recode(
    measure = c("GVAHRS"),
    activity = c("Total economy" = "_T", "Private sector" = "BTNXL"),
    vars = c("fp_2020_ppp17")
  ) |> 
  mutate(values = rebase(values, time, baseyear = 2007), .by = where(is.factor)) |> 
  filter(geo != "DK") |> 
  filter_recode(
    geo = geos
  ) |> 

  ggplot(aes(time, values, colour = geo, linewidth = geo)) +
  facet_wrap(~ activity, nrow = 1) +
  geom_line() +
  scale_colour_manual(values = geos_colour) +
  scale_linewidth_manual(values = geos_width) +
  guides(colour = guide_legend(reverse = TRUE), linewidth = guide_legend(reverse = TRUE)) +
  the_title_blank("xyl") +
  labs(
    title = "Labour productivity, value added / hours worked",
    subtitle = "Index, 2007 = 100",
    caption = "Source: OECD, Finnish Productivity Board"
  )
```

### Industries outside the private sector

```{r}

dat_gva_ind |> 
  filter(time >= "1995-01-01") |> 
  filter_recode(
    measure = c("GVAHRS"),
    activity = c("Real estate" = "L", "Public sector" = "OTQ", "Other services" = "RTU"),
    vars = c("fp_2020_ppp17")
  ) |> 
  mutate(values = rebase(values, time, baseyear = 2007), .by = where(is.factor)) |> 
  filter(geo != "DK") |> 
  filter_recode(
    geo = geos
  ) |> 

  ggplot(aes(time, values, colour = geo, linewidth = geo)) +
  facet_wrap(~ activity, nrow = 1) +
  geom_line() +
  scale_colour_manual(values = geos_colour) +
  scale_linewidth_manual(values = geos_width) +
  guides(colour = guide_legend(reverse = TRUE), linewidth = guide_legend(reverse = TRUE)) +
  the_title_blank("xyl") +
  labs(
    title = "Labour productivity, value added / hours worked",
    subtitle = "Index, 2007 = 100",
    caption = "Source: OECD, Finnish Productivity Board"
  )
```

### Private sector breakdown

```{r}

dat_gva_ind|> 
  filter(time >= "1995-01-01") |> 
  filter_recode(
    measure = c("GVAHRS"),
    activity = c("Industry" = "BTE", "Construction" = "F","Private services" = "GTNXL"),
    vars = c("fp_2020_ppp17")
  ) |> 
  mutate(values = rebase(values, time, baseyear = 2007), .by = where(is.factor)) |> 
  filter(geo != "DK") |> 
  filter_recode(
    geo = geos
  ) |> 

  ggplot(aes(time, values, colour = geo, linewidth = geo)) +
  scale_y_continuous(limits = c(50, 150)) +
  facet_wrap(~ activity) +
  geom_line() +
  scale_colour_manual(values = geos_colour) +
  scale_linewidth_manual(values = geos_width) +
  guides(colour = guide_legend(reverse = TRUE), linewidth = guide_legend(reverse = TRUE)) +
  the_title_blank("xyl") +
  labs(
    title = "Labour productivity, value added / hours worked",
    subtitle = "Index, 2007 = 100",
    caption = "Source: OECD, Finnish Productivity Board"
  )
```

### Shares

```{r}

dat_gva_ind |> 
  filter(time >= "1995-01-01") |> 
  filter_recode(
    measure = c("GVA"),
    activity = c("_T", "BTNXL", "L", "OTQ", "RTU"),
    vars = c("cp")
  ) |> 
  mutate(values = 100 * values / values[activity == "_T"] , .by = !c(activity, values)) |> 
  filter_recode(
    geo = geos,
    activity = c("Private sector" = "BTNXL", "Real estate" = "L", "Public sector" = "OTQ", "Other services" = "RTU")) |> 
  ggplot(aes(time, values, colour = activity)) +
  facet_wrap(~ geo, nrow = 1) +
  geom_line() +
  scale_x_date(date_labels = "%y") +
  the_title_blank("xyl") +
  labs(
    title = "Share of value added",
    subtitle = "%",
    caption = "Source: OECD, Finnish Productivity Board"
  ) 

```

```{r}

dat_gva_ind |> 
  filter(time >= "1995-01-01") |> 
  filter_recode(
    measure = c("GVA"),
    activity = c("_T", "A", "BTE", "F", "GTNXL", "L", "OTQ", "RTU"),
    vars = c("cp")
  ) |> 
  spread(activity, values) |> 
  rowwise() |> 
  mutate(Other = `_T` - sum(c_across(all_of(c("A", "BTE", "F", "GTNXL", "L", "OTQ", "RTU"))), na.rm = TRUE),
         Other = if_else(Other == `_T`, NA, Other)) |> 
  ungroup() |> 
  pivot_longer(where(is.numeric), names_to = "activity", values_to = "values") |> 
  mutate(values = 100 * values / values[activity == "_T"] , .by = !c(activity, values)) |> 
  filter_recode(
    geo = geos,
    activity = c("Industry" = "BTE", "Construction" = "F","Private services" = "GTNXL", "Real estate" = "L", "Public-type services" = "OTQ", "Other services" = "RTU", "Agriculture" = "A", "Other")) |> 
  ggplot(aes(time, values, colour = activity)) +
  facet_wrap(~ geo, nrow = 1) +
  geom_line() +
  scale_x_date(date_labels = "%y") +
  the_title_blank("xyl") +
  labs(
    title = "Share of value added",
    subtitle = "%",
    caption = "Source: OECD, Finnish Productivity Board"
  ) 

```

There is an error in OECD data. Other should be zero, but in the data Industry includes only manufacturing.

```{r}

dat_gva_ind |> 
  filter(time >= "1995-01-01") |> 
  filter_recode(
    measure = c("GVA", "GVAHRS"),
    activity = c("_T", "BTE", "F", "GTNXL", "BTNXL", "L", "OTQ", "RTU", "A"),
    vars = c("cp")
  ) |> 
  spread(activity, values) |> 
  rowwise() |> 
  mutate(Other = `_T` - sum(c_across(all_of(c("A", "BTE", "F", "GTNXL", "L", "OTQ", "RTU"))), na.rm = TRUE),
         Other = if_else(Other == `_T`, NA, Other)) |> 
  ungroup() |> 
  pivot_longer(where(is.numeric), names_to = "activity", values_to = "values") |> 
  spread(measure, values) |> 
  mutate(HRS = GVA / GVAHRS) |> 
  select(-GVA, -GVAHRS) |> 
  pivot_longer(HRS, names_to = "measure", values_to = "values") |> 
  mutate(values = 100 * values / values[activity == "_T"] , .by = !c(activity, values)) |> 
  filter_recode(
    geo = geos,
activity = c("Industry" = "BTE", "Construction" = "F","Private services" = "GTNXL", "Real estate" = "L", "Public-type services" = "OTQ", "Other services" = "RTU", "Agriculture" = "A", "Other")) |> 
  ggplot(aes(time, values, colour = activity)) +
  facet_wrap(~ geo, nrow = 1) +
  geom_line() +
  scale_x_date(date_labels = "%y") +
  the_title_blank("xyl") +
  labs(
    title = "Share of hours worked",
    subtitle = "%",
    caption = "Source: OECD, Finnish Productivity Board"
  ) 

```

## GDP and productivity levels

GDP and productivity growth is easier to compare than levels, because countries have different price levels and exchange rates also vary a lot.

```{r}

dat_oecd_pdb_main |> 
  filter(time >= "1995-01-01") |> 
  filter_recode(
    measure = c("GDPPOP"),
    activity = c("Total economy" = "_T"),
    price_base = c("LR"),        # 
    conversion_type = c("PPP")
  ) |> 
  mutate(geo2 = suppressWarnings(fct_relevel(geo, geos, after = Inf))) |>
  mutate(geo = fct_other(geo, keep = geos, other_level = "Other")) |> 
  filter_recode(geo = geos) |> 
  select(-unit_measure) |> 
  ggplot(aes(time, values, group = geo2, colour = geo, linewidth = geo)) +
  geom_line() +
  scale_colour_manual(values = geos_colour) +
  scale_linewidth_manual(values = geos_width) +
  guides(colour = guide_legend(reverse = TRUE), linewidth = guide_legend(reverse = TRUE)) +
  the_title_blank("xyl") +
  labs(
    title = "GDP relative to population",
    subtitle = "At 2020 prices, PPP-adjusted",
    caption = "Source: OECD, Finnish Productivity Board"
  )

# ggptt::ggsave_ppt_half("bkt_capita", plot = last_plot() filter(p@data, geo != "Denmark") +theme_vm() +the_title_blank("xyl"))

# last_plot()$data |> filter(geo != "others")  |> select(geo, time, values) |> spread(geo, values) |> conc()

```

### At constant prices with exchange rates and PPP

Base year 2020

```{r}

dat_oecd_pdb_main |> 
  filter(time >= "1995-01-01") |> 
  filter_recode(
    measure = c("GDPPOP"),
    activity = c("Total economy" = "_T"),
    price_base = c("LR"),        # 
    conversion_type = c(MP = "_Z","PPP")
  ) |> 
  select(-unit_measure,  -var_id) |> 
  spread(conversion_type, values) |>
  # For some reason, OECD data has no market-price series for the USA, but it equals PPP because PPP is relative to the USA.
  mutate(MP = if_else(geo == "US", PPP, MP)) |>
  mutate(MP = convert_currency(MP, geo, time, to = "USD", base_time = "2020-01-01")) |> 
  pivot_longer(where(is.numeric), names_to = "conversion_type", values_to = "values") |>
  filter_recode(
    geo = geos,
    conversion_type = c("Exchange rate" = "MP", "Purchasing power parity" ="PPP")
  ) |>
  ggplot(aes(time, values, colour = geo, linewidth = geo)) +
  facet_wrap(~ conversion_type) +
  geom_line() +
  scale_colour_manual(values = geos_colour) +
  scale_linewidth_manual(values = geos_width) +
  guides(colour = guide_legend(reverse = TRUE), linewidth = guide_legend(reverse = TRUE)) +
  the_title_blank("xyl") +
  labs(
    title = "GDP per capita at constant prices",
    subtitle = "2020 USD",
    caption = "Source: OECD, Finnish Productivity Board"
  )
  

```

### At current prices with exchange rates and PPP

```{r}

dat_oecd_pdb_main |> 
  filter(time >= "1995-01-01") |> 
  filter_recode(
     measure = c("GDPPOP"),
    activity = c("Total economy" = "_T"),
    price_base = c("V"),        # 
    conversion_type = c(MP = "_Z","PPP")
  ) |> 
  select(-unit_measure,  -var_id) |> 
  spread(conversion_type, values) |> 
  # For some reason, OECD data has no market-price series for the USA, but it equals PPP because PPP is relative to the USA.
  # mutate(MP = if_else(geo == "US", PPP, MP)) |> 
  mutate(MP1 = convert_currency(MP, geo, time, to = "USD", base_time = "2020-01-01")) |>
  mutate(MP2 = convert_currency(MP, geo, time, to = "USD")) |> 
  select(-MP) |> 
  pivot_longer(where(is.numeric), names_to = "conversion_type", values_to = "values") |> 
  filter_recode(
    geo = geos,
    conversion_type = c("Exchange rate, variable" = "MP2","Exchange rate 2020" = "MP1" , "Purchasing power parity 2020" ="PPP")
  ) |>
  ggplot(aes(time, values, colour = geo, linewidth = geo)) +
  facet_wrap(~ conversion_type) +
  geom_line() +
  scale_colour_manual(values = geos_colour) +
  scale_linewidth_manual(values = geos_width) +
  guides(colour = guide_legend(reverse = TRUE), linewidth = guide_legend(reverse = TRUE)) +
  the_title_blank("xyl") +
  labs(
    title = "GDP per capita at current prices",
    subtitle = "2020 USD",
    caption = "Source: OECD, Finnish Productivity Board"
  )

```

## Level gap in productivity and hours worked

Is the GDP level gap explained by productivity or by hours worked?

In the OECD database, labour productivity is measured with value added per hour worked, and hours with hours per capita.

Clearly, the level gap comes from labour productivity, not from hours worked. However, the labour productivity level gap to the United States is much larger than GDP per capita, even though hours per capita are almost the same. Why?

```{r}

dat_oecd_pdb_main |> 
  filter(time >= "1995-01-01") |> 
  filter_recode(
    geo = geos,
    measure = c("Labour productivity" = "GVAHRS", "Hours per capita" = "HRSPOP"),
    activity = c("Total economy" = "_T"),
    unit_measure = c("USD_PPP_H", "H_PS"),
    price_base = c("LR", "_Z"),        
    conversion_type = c("PPP", "_Z")
  ) |> 
  select(-unit_measure) |> 
  ggplot(aes(time, values, colour = geo, linewidth = geo)) +
  facet_wrap(~measure, scales = "free") +
  geom_line() +
  scale_colour_manual(values = geos_colour) +
  scale_linewidth_manual(values = geos_width) +
  guides(colour = guide_legend(reverse = TRUE), linewidth = guide_legend(reverse = TRUE)) +
  the_title_blank("xyl") +
  labs(
    title = "Labour productivity and hours worked",
    subtitle = "Index, 2007 = 100",
    caption = "Source: OECD, Finnish Productivity Board"
  )

```

The reason is different statistical practices between the United States and Eurostat. GDP per hour worked shows differences similar to GDP per capita. Value added per hour worked is instead clearly higher in the United States than in EU countries.

In European national accounts, GDP = value added + product taxes - subsidies. In the United States, GDP and value added are the same, and product taxes and subsidies are already included at the value-added level.

```{r}

dat_oecd_pdb_main |> 
  filter(time >= "1995-01-01") |> 
  filter(var_id %in% c("GVAHRS-USD_PPP_H-LR-PPP", 
                       "GDP-USD_PPP-LR-PPP", 
                       "GVA-USD_PPP-LR-PPP", 
                       "HRSPOP-H_PS-_Z-_Z")) |> 
  filter_recode(
    geo = geos,
    measure = c("GVAHRS", "GDP", "GVA", "HRSPOP"),
    activity = c("Total economy" = "_T"),
    unit_measure = c("USD_PPP_H", "USD_PPP", "H_PS"),
    price_base = c("LR", "_Z"),        
    conversion_type = c("PPP", "_Z")
  ) |> 
  select(time, geo, measure, values) |> 
  spread(measure, values) |> 
  mutate(HRS = GVA / GVAHRS,
         GDPHRS = GDP / HRS) |> 
  pivot_longer(where(is.numeric), names_to = "measure", values_to = "values") |> 
  filter_recode(
    measure = c("GDPHRS", "GVAHRS")
  ) |> 
  ggplot(aes(time, values, colour = geo, linewidth = geo)) +
  facet_wrap(~measure, scales = "free") +
  geom_line() +
  scale_colour_manual(values = geos_colour) +
  scale_linewidth_manual(values = geos_width) +
  guides(colour = guide_legend(reverse = TRUE), linewidth = guide_legend(reverse = TRUE)) +
  the_title_blank("xyl") +
  labs(
    title = "GDP and value added per hour worked",
    subtitle = "PPP-adjusted, 2020 $",
    caption = "Source: OECD, Finnish Productivity Board"
  )

```

```{r}

dat_oecd_pdb_main |> 
  filter(time >= "1995-01-01") |> 
  filter_recode(
    geo = geos,
    measure = c("GDP", "GVA"),
    activity = c("Total economy" = "_T"),
    # unit_measure = c("USD_PPP"),
    price_base = c("V") ,        
    conversion_type = c("PPP")
  ) |> 
  select(-unit_measure) |> 
  ggplot(aes(time, values, colour = measure)) +
  facet_wrap(~geo, scales = "free") +
  geom_line() +
  the_title_blank("xyl") +
  labs(
    title = "GDP and value added",
    subtitle = "PPP-adjusted, 2020 $",
    caption = "Source: OECD, Finnish Productivity Board"
  ) +
  geom_hline(yintercept = 0)

```

GDP can only be used for the whole economy. Value added must be used when the economy is examined by industries or industry groups.

Productivity differences in the private sector are not as large, even when value added is used.

In OECD statistics, the private sector means the part of the economy that can be measured reasonably reliably, including broad industry (B, C, E, D), construction (F), and market services (G-N excl. L). It excludes agriculture, real estate, public-type services (public administration, education, and health and social services; these also include private services), and small service industries (R, S, T).

```{r}

dat_oecd_pdb_main |> 
  filter(time >= "1995-01-01") |> 
  filter_recode(
    measure = c("GVAHRS"),
    activity = c("Total economy" = "_T", "Private" = "BTNXL"),
    price_base = c("LR"),        # 
    conversion_type = c("_Z")
  ) |> 
  # group_by(across(where(is.factor))) |> 
  # mutate(values = rebase(values, time, 2005)) |> 
  # ungroup() |> |> 
  mutate(values = convert_currency(values, geo, time, to = "USD", base_time = "2020-01-01")) |>   filter_recode(
    geo = geos
  )|> 
  ggplot(aes(time, values, colour = geo, linewidth = geo)) +
  facet_wrap(~ activity) +
  geom_line() +
  scale_colour_manual(values = geos_colour) +
  scale_linewidth_manual(values = geos_width) +
  guides(colour = guide_legend(reverse = TRUE), linewidth = guide_legend(reverse = TRUE)) +
  the_title_blank("xyl") +
  labs(
    title = "Value added / hours worked",
    subtitle = "2020 $",
    caption = "Source: OECD, Finnish Productivity Board"
  ) 

```

## OECD and GGDC difference in PPP

Industry-level data are price-level adjusted using industry PPP price levels from the GGDC Productivity Level Database. The OECD database does not include PPP-adjusted series for industries.

Prices for the total economy and industry groups are aggregated as value-added-weighted averages. Below, differences in price-level-adjusted gross value added at the total-economy level are examined using OECD and GGDC PPP data.

A small difference between series is expected, because OECD uses base year 2020 and GGDC uses 2017. For the countries examined, the differences are explained by the base year.

```{r}
# 
# dat_gva_ind |> str()
# dat_oecd_pdb_main |> str()

dat1 <- dat_oecd_pdb_main |> 
  filter(time >= "1995-01-01") |> 
  filter_recode(
    geo = geos,
    measure = c("GVA"),
    activity = c("_T"),
    price_base = c("LR"),        # 
    conversion_type = c("PPP")
  ) |> 
  select(time, geo, values) |> 
  mutate(ppp_source = "oecd")
  
dat2 <- dat_gva_ind |> 
  filter(time >= "1995-01-01") |> 
  filter_recode(
    geo = geos,
    measure = c("GVA"),
    activity = c("_T"),
    vars = "fp_2020_ppp17"
  ) |> 
  select(time, geo, values) |> 
  mutate(ppp_source = "ggdc")

bind_rows(dat1, dat2) |> 
  mutate(values = values / 1000) |> 
  ggplot(aes(time, values, colour = ppp_source)) +
  facet_wrap(~geo, scales = "free") +
  geom_line() +
  the_title_blank("xyl") +
  the_legend_bot() +
  labs(title = "Gross value added in the total economy",
       subtitle = "Billion PPP dollars at 2020 or 2017 prices")

```

```{r}
#| fig-width: 7
#| fig-height: 7

dat_gva_ind|> 
  filter(time >= "1995-01-01") |> 
  filter_recode(
    measure = c("GVAHRS"),
    activity = c("Total economy" = "_T", "Private" = "BTNXL", "Public-type" = "OTQ"),
    vars = c("Exchange rate" = "fp_2020_xr17", "PPP" = "fp_2020_ppp17")
  ) |> 
  # mutate(values = rebase(values, time, baseyear = 2007), .by = where(is.factor)) |> 
  filter_recode(
    geo = geos
  ) |> 
  ggplot(aes(time, values, colour = geo, linewidth = geo)) +
  # scale_y_continuous(limits = c(50, 150)) +
  facet_grid(vars ~ activity) +
  geom_line() +
  scale_colour_manual(values = geos_colour) +
  scale_linewidth_manual(values = geos_width) +
  guides(colour = guide_legend(reverse = TRUE), linewidth = guide_legend(reverse = TRUE)) +
  the_title_blank("xyl") +
  labs(
    title = "Value added / hours worked",
    subtitle = "2017 $, with exchange rates and PPP adjustment",
    caption = "Source: OECD, Finnish Productivity Board"
  )

```

```{r}
#| fig-width: 7
#| fig-height: 7
dat_gva_ind|> 
  filter(time >= "1995-01-01") |> 
  filter_recode(
    measure = c("GVAHRS"),
    activity = c("Real estate" = "L", "Other services" = "RTU"),
    vars = c("Exchange rate" = "fp_2020_xr17", "PPP" = "fp_2020_ppp17")
  ) |> 
  # mutate(values = rebase(values, time, baseyear = 2007), .by = where(is.factor)) |> 
  filter_recode(
    geo = geos
  ) |> 
  ggplot(aes(time, values, colour = geo, linewidth = geo)) +
  # scale_y_continuous(limits = c(50, 150)) +
  facet_wrap(vars ~ activity, scales = "free") +
  geom_line() +
  scale_colour_manual(values = geos_colour) +
  scale_linewidth_manual(values = geos_width) +
  guides(colour = guide_legend(reverse = TRUE), linewidth = guide_legend(reverse = TRUE)) +
  the_title_blank("xyl") +
  labs(
    title = "Value added / hours worked",
    subtitle = "2017 $, with exchange rates and PPP adjustment",
    caption = "Source: OECD, Finnish Productivity Board"
  ) 

```

```{r}

dat_gva_ind|> 
  filter(time >= "1995-01-01") |> 
  filter_recode(
    measure = c("GVAHRS"),
    activity = c("Industry" = "BTE","Private services" = "GTNXL"),
    vars = c("fp_2020_ppp17")
  ) |> 
  # mutate(values = rebase(values, time, baseyear = 2007), .by = where(is.factor)) |> 
  filter_recode(
    geo = geos
  ) |> 
  ggplot(aes(time, values, colour = geo, linewidth = geo)) +
  # scale_y_continuous(limits = c(50, 150)) +
  facet_wrap(~ activity) +
  geom_line() +
  scale_colour_manual(values = geos_colour) +
  scale_linewidth_manual(values = geos_width) +
  guides(colour = guide_legend(reverse = TRUE), linewidth = guide_legend(reverse = TRUE)) +
  the_title_blank("xyl") +
  labs(
    title = "Value added / hours worked",
    subtitle = "2017 $, PPP-adjusted",
    caption = "Source: OECD, Finnish Productivity Board"
  ) 

```

```{r}
#| fig-width: 7
#| fig-height: 7

dat_gva_ind|> 
  filter(time >= "1995-01-01") |> 
  filter_recode(
    measure = c("GVAHRS"),
    activity = c("Industry" = "BTE", "Construction" = "F", "Private services" = "GTNXL"),
    vars = c("Exchange rate" = "fp_2020_xr17", "PPP" = "fp_2020_ppp17")
  ) |> 
  # mutate(values = rebase(values, time, baseyear = 2007), .by = where(is.factor)) |> 
  filter_recode(
    geo = geos
  ) |> 
  ggplot(aes(time, values, colour = geo, linewidth = geo)) +
  # scale_y_continuous(limits = c(50, 150)) +
  facet_grid(vars ~ activity) +
  geom_line() +
  scale_colour_manual(values = geos_colour) +
  scale_linewidth_manual(values = geos_width) +
  guides(colour = guide_legend(reverse = TRUE), linewidth = guide_legend(reverse = TRUE)) +
  the_title_blank("xyl") +
  labs(
    title = "Value added / hours worked",
    subtitle = "2017 $, PPP-adjusted",
    caption = "Source: OECD, Finnish Productivity Board"
  ) 

```

```{r}
#| fig-width: 7
#| fig-height: 7

dat_gva_ind|> 
  filter(time >= "1995-01-01") |> 
  filter_recode(
    measure = c("GVAHRS"),
    activity = c("Industry" = "BTE", "Private services" = "GTNXL"),
    vars = c("Exchange rate" = "fp_2020_xr17", "PPP" = "fp_2020_ppp17")
  ) |> 
  # mutate(values = rebase(values, time, baseyear = 2007), .by = where(is.factor)) |> 
  filter_recode(
    geo = geos
  ) |> 
  ggplot(aes(time, values, colour = geo, linewidth = geo)) +
  # scale_y_continuous(limits = c(50, 150)) +
  facet_grid(vars ~ activity) +
  geom_line() +
  scale_colour_manual(values = geos_colour) +
  scale_linewidth_manual(values = geos_width) +
  guides(colour = guide_legend(reverse = TRUE), linewidth = guide_legend(reverse = TRUE)) +
  the_title_blank("xyl") +
  labs(
    title = "Value added / hours worked",
    subtitle = "2017 $, PPP-adjusted",
    caption = "Source: OECD, Finnish Productivity Board"
  ) 

```
